---
title: "dplyr"
author: "Xiaoqian Xue"
date: "September 22, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```

## Data Manipulation(1)
Pick observations by their values (filter()). 
```{r}
filter(flights)
```

## Data Manipulation(2)
Reorder the rows (arrange()).
```{r}
arrange(flights,desc(dep_time))
```

## Data Manipulation(3)
Pick variables by their names (select()).  
```{r}
select(flights, dep_time)
```

## Data Manipulation(4)
Create new variables with functions of existing variables (mutate()).
```{r}
mutate(flights,speed = distance/air_time)
```

## Data Manipulation(5)
Collapse many values down to a single summary (summarise()).
```{r}
summarise(flights)
```

## All verbs work similarly:  

The first argument is a data frame.  

The subsequent arguments describe what to do with the data frame, using the variable names (without quotes).  

The result is a new data frame.  

## OK, but what is a tibble?

[Look here for everything you want to know](http://bit.ly/2ffxvfM)  


Moving on with dplyr ...  

## filter()

filter() allows you to subset observations based on their values.

```{r}
filter(flights, month == 1, day == 1)

```


## R either prints out the results, or saves them to a variable.

If you want to do both, you can wrap the assignment in parentheses:

```{r}
(jan1 <- filter(flights, month == 1, day == 1))

```

## NA -- key knowledge for statisticians

```{r}
NA > 5

10 == NA

NA + 10

NA / 2

## -------------------

NA == NA

```

## arrange()

arrange() works similarly to filter()   
except that instead of selecting rows,   
it changes their order.

```{r}
arrange(flights, year, month, day)
```


## desc()

Use desc() to re-order by a column in descending order:

```{r}
arrange(flights, desc(arr_delay))
```




